---
title: "Monitoring Manager Assignment"
author: "Tasbiul"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
geometry: margin=0.75in
fontsize: 10pt
header-includes:
  - |
    \usepackage{xcolor}
    \definecolor{mainpink}{HTML}{C2185B}
    \definecolor{subpink}{HTML}{C85C78}
    \newenvironment{mainquestion}{\color{mainpink}\bfseries}{}
    \newenvironment{subquestion}{\color{subpink}\bfseries}{}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 7.2,
  fig.height = 4.2,
  dpi = 150
)

library(dplyr)
library(ggplot2)
library(knitr)
library(readxl)
library(scales)
library(stringr)
library(writexl)

options(dplyr.summarise.inform = FALSE, scipen = 999)
```

<style>
.mainquestion {
  color: #C2185B;
  background: #FCE4EC;
  border-left: 5px solid #C2185B;
  padding: 10px 14px;
  margin: 14px 0;
  font-weight: 700;
}
.subquestion {
  color: #B94F6D;
  background: #FFF1F4;
  border-left: 4px solid #D98298;
  padding: 8px 12px;
  margin: 12px 0;
  font-weight: 600;
}
</style>

# Scope and method

I used R so the work can be checked, repeated, and updated if the source files change.

**Data sources:**

- `Average_sessions_details.xlsx`
- `CCP_Monitoring_Data.xlsx`

**Assignment instructions:**

- `Monitoring Manager Assignment.pdf`

```{r import-functions}
# Purpose: standardize text before comparing or joining names.
# Input: any text column (for example, district or facility names).
# Output: the same text with consistent apostrophes and extra spaces removed.
clean_text <- function(x) {
  x |>
    as.character() |>               # Ensure every value is treated as text.
    str_replace_all("[‘’]", "'") |>  # Change smart apostrophes to a standard apostrophe.
    str_squish()                      # Remove leading, trailing, and repeated spaces.
}

# Purpose: create a comparison-only version of a name to find hidden duplicates.
# Input: cleaned or uncleaned names.
# Output: lowercase letters/numbers only; this key is never shown as the final name.
string_key <- function(x) {
  x |>
    clean_text() |>                       # Apply the basic text cleanup above.
    iconv(to = "ASCII//TRANSLIT") |>        # Convert unusual characters to plain text.
    str_to_lower() |>                        # Ignore capitalization differences.
    str_replace_all("[^a-z0-9]", "")           # Ignore punctuation and spaces when comparing.
}

# Reviewed spelling map: left side is the source variation; right side is the chosen name.
district_name_crosswalk <- c(
  "Bogra" = "Bogura",
  "Jhalakathi" = "Jhalokathi",
  "Laxmipur" = "Lakshmipur",
  "Narshingdi" = "Narsingdi",
  "Sariatpur" = "Shariatpur",
  "Thakurgaong" = "Thakurgaon"
)

# The same reviewed spelling corrections when a district appears inside a longer facility name.
# In these patterns, `\\b` means "whole word" so part of another word is not replaced.
facility_location_patterns <- c(
  "\\bBogra\\b" = "Bogura", "\\bJhalakathi\\b" = "Jhalokathi",
  "\\bLaxmipur\\b" = "Lakshmipur", "\\bNarshingdi\\b" = "Narsingdi",
  "\\bSariatpur\\b" = "Shariatpur", "\\bThakurgaong\\b" = "Thakurgaon",
  "\\bKhagrachari\\b" = "Khagrachhari", "\\bBola\\b" = "Bhola",
  "\\bMoulavibazar\\b" = "Moulvibazar", "\\bKustia\\b" = "Kushtia",
  "\\bChaattogram\\b" = "Chattogram", "\\bLaksmipur\\b" = "Lakshmipur",
  "\\bPatukhali\\b" = "Patuakhali", "\\bSulhet\\b" = "Sylhet"
)

# Purpose: flag unusually high or low values for review without changing them.
# Input: a numeric column and a multiplier; 3 gives a conservative review threshold.
# Output: TRUE for unusual observations and FALSE for observations within the range.
iqr_outlier <- function(x, multiplier = 3) {
  quartiles <- quantile(x, c(0.25, 0.75), na.rm = TRUE) # Middle 50% boundaries.
  spread <- diff(quartiles)                             # Interquartile range (IQR).

  # If every value is effectively the same, none can be classified as unusual.
  if (!is.finite(spread) || spread == 0) return(rep(FALSE, length(x)))

  # Return flags only; the original values remain unchanged.
  x < quartiles[1] - multiplier * spread |
    x > quartiles[2] + multiplier * spread
}
```

# Data problems and fixes

```{r import-and-clean}
monitor_raw <- read_excel("CCP_Monitoring_Data.xlsx")
history_raw <- read_excel("Average_sessions_details.xlsx")

monitor_raw_names <- names(monitor_raw)

monitor <- monitor_raw
history <- history_raw

# Issue: one source header says "Surgury"; rename fields without changing values.
monitor <- monitor |>
  rename(
    division = Division,
    district = District,
    facilities = Facilities,
    anc_sessions = `ANC Sessions`,
    pnc_sessions = `PNC Sessions`,
    scanu_sessions = `SCANU Sessions`,
    general_medicine_sessions = `General Medicine Sessions`,
    general_surgery_sessions = `General Surgury Sessions`,
    ncd_sessions = `NCD Sessions`,
    total_patients = `Total Patients`,
    total_caregivers = `Total Caregivers`,
    trainers = Trainers,
    avg_sessions_week = `Avg Sessions/Week`,
    avg_participants_session = `Avg Participants/Session`,
    sessions_vs_target_pct = `% Sessions vs Target`,
    ccp_quality_assessment_coverage_pct = `CCP Quality Assessment Coverage %`
  )

history <- history |>
  rename(
    active_since = `Active since`,
    division = Division,
    district = District,
    facility_type = `Facility Type`,
    facility_name = `Facility name`,
    total_number_of_sessions = `Total number of sessions`,
    average_monthly_sessions = `Average monthly sessions`,
    total_number_of_participants = `Total number of participants`,
    average_monthly_participants = `Average monthly participants`
  )

required_monitor <- c(
  "division", "district", "facilities", "anc_sessions", "pnc_sessions",
  "scanu_sessions", "general_medicine_sessions", "general_surgery_sessions",
  "ncd_sessions", "total_patients", "total_caregivers", "trainers",
  "avg_sessions_week", "avg_participants_session", "sessions_vs_target_pct",
  "ccp_quality_assessment_coverage_pct"
)
required_history <- c(
  "active_since", "division", "district", "facility_type", "facility_name",
  "total_number_of_sessions", "average_monthly_sessions",
  "total_number_of_participants", "average_monthly_participants"
)

stopifnot(all(required_monitor %in% names(monitor)))
stopifnot(all(required_history %in% names(history)))

monitor_numeric <- setdiff(required_monitor, c("division", "district"))
history_numeric <- c(
  "active_since", "total_number_of_sessions", "average_monthly_sessions",
  "total_number_of_participants", "average_monthly_participants"
)

# readxl reads each column type; stop with a clear message if a number column contains text.
numeric_type_failures <-
  sum(!vapply(monitor[monitor_numeric], is.numeric, logical(1))) +
  sum(!vapply(history[history_numeric], is.numeric, logical(1)))
stopifnot(numeric_type_failures == 0)

monitor <- monitor |>
  mutate(
    source_row = row_number() + 1L,
    division_original = as.character(division),
    district_original = as.character(district),
    division_raw = clean_text(division_original),
    district_raw = clean_text(district_original),
    division = recode(division_raw, "Barisal" = "Barishal"),
    district = recode(district_raw, !!!as.list(district_name_crosswalk))
  )

history <- history |>
  mutate(
    source_row = row_number() + 1L,
    division_original = as.character(division),
    district_original = as.character(district),
    facility_type_original = as.character(facility_type),
    facility_name_original = as.character(facility_name),
    division_raw = clean_text(division_original),
    district_raw = clean_text(district_original),
    facility_type = clean_text(facility_type),
    facility_name = clean_text(facility_name_original) |>
      str_replace_all("\\s*,+\\s*", ", ") |>
      str_replace_all(facility_location_patterns) |>
      str_remove(",\\s*$") |>
      str_squish(),
    division = recode(division_raw, "Barisal" = "Barishal"),
    district = recode(district_raw, !!!as.list(district_name_crosswalk))
  )

# Issue: curly apostrophes and spelling differences stop the same place names from matching.
# Fix: make punctuation consistent and correct only the known place-name differences above.

# Issue: facility names contain line breaks, extra spaces, misplaced commas, and place-name typos.
# Fix: tidy the spaces and commas, then correct only the place names checked above.
facility_name_fixes <- history |>
  filter(facility_name_original != facility_name) |>
  transmute(
    row = source_row,
    original = facility_name_original,
    corrected = facility_name
  )

session_fields <- c(
  "anc_sessions", "pnc_sessions", "scanu_sessions",
  "general_medicine_sessions", "general_surgery_sessions", "ncd_sessions"
)

monitor <- monitor |>
  mutate(
    total_sessions_quarter = rowSums(across(all_of(session_fields)), na.rm = FALSE),
    total_participants = total_patients + total_caregivers,
    calculated_sessions_week = total_sessions_quarter / 13,
    calculated_participants_session = total_participants / total_sessions_quarter
  )

# Issue: four monitoring rows use city units, while the older file uses their parent districts.
# Fix: keep the original district and use a second district name only for a fair comparison.
monitor <- monitor |>
  mutate(
    comparison_district = case_when(
      district == "Chattogram City" ~ "Chattogram",
      district %in% c("Dhaka North City", "Dhaka South City") ~ "Dhaka",
      district == "Khulna City" ~ "Khulna",
      TRUE ~ district
    )
  )

history <- history |>
  mutate(
    comparison_district = district,
    implied_months_sessions = total_number_of_sessions / average_monthly_sessions,
    implied_months_participants = total_number_of_participants / average_monthly_participants,
    calculated_participants_session = total_number_of_participants / total_number_of_sessions,
    average_participants_session = average_monthly_participants / average_monthly_sessions
  ) |>
  group_by(facility_type) |>
  mutate(
    unusual_monthly_sessions = iqr_outlier(average_monthly_sessions),
    unusual_participants_session = iqr_outlier(calculated_participants_session)
  ) |>
  ungroup()

monitor <- monitor |>
  group_by(division) |>
  mutate(
    unusual_target_pct = iqr_outlier(sessions_vs_target_pct),
    unusual_qa_pct = iqr_outlier(ccp_quality_assessment_coverage_pct),
    unusual_sessions_per_facility = iqr_outlier(total_sessions_quarter / facilities),
    unusual_participants_session = iqr_outlier(calculated_participants_session)
  ) |>
  ungroup()

# Issue: a very high or low value may be a mistake, but it may also be real.
# Fix: mark values outside the 3-IQR range for checking and do not change them.
```

::: {.mainquestion latex=true}
**Question:** What problems are in the two files, and what did I do about them?
:::

**Table: Problems found and how they were handled**

```{r validation-audit}
monitor_district <- monitor |>
  group_by(division, comparison_district) |>
  summarise(
    monitoring_rows = n(),
    facilities_monitoring = sum(facilities),
    reported_quarter_sessions = sum(total_sessions_quarter),
    reported_quarter_participants = sum(total_participants),
    .groups = "drop"
  )

history_district <- history |>
  group_by(division, comparison_district) |>
  summarise(
    facilities_history = n_distinct(facility_name),
    expected_quarter_sessions = sum(average_monthly_sessions) * 3,
    expected_quarter_participants = sum(average_monthly_participants) * 3,
    .groups = "drop"
  )

district_comparison_all <- full_join(
  monitor_district,
  history_district,
  by = c("division", "comparison_district")
) |>
  mutate(
    join_status = case_when(
      is.na(reported_quarter_sessions) ~ "History only",
      is.na(expected_quarter_sessions) ~ "Monitoring only",
      TRUE ~ "Matched"
    ),
    facility_count_gap = facilities_monitoring - facilities_history,
    session_gap = reported_quarter_sessions - expected_quarter_sessions,
    session_gap_pct = session_gap / expected_quarter_sessions
  )

# Issue: both files total 658 facilities, but most district counts differ.
# Fix: do not move records without a dated matching list; show every difference.

# Issue: the brief says 8 facility categories, while the workbook contains 9.
# Fix: keep all 9 categories found in the file and note that the brief says 8.

# Issue: only the start year is given, and the two files may cover different periods.
# Fix: use monthly average x 3 only as a rough comparison and clearly state this limit.

monitor_count_fields <- c(
  "facilities", session_fields, "total_patients", "total_caregivers", "trainers"
)
history_count_fields <- c(
  "active_since", "total_number_of_sessions", "total_number_of_participants"
)

non_finite_numeric <-
  sum(is.infinite(unlist(monitor[monitor_numeric])) |
        is.nan(unlist(monitor[monitor_numeric])), na.rm = TRUE) +
  sum(is.infinite(unlist(history[history_numeric])) |
        is.nan(unlist(history[history_numeric])), na.rm = TRUE)

non_positive_counts <-
  sum(unlist(monitor[monitor_count_fields]) <= 0, na.rm = TRUE) +
  sum(unlist(history[setdiff(history_count_fields, "active_since")]) <= 0, na.rm = TRUE)

non_integer_counts <-
  sum(abs(unlist(monitor[monitor_count_fields]) -
            round(unlist(monitor[monitor_count_fields]))) > 1e-9, na.rm = TRUE) +
  sum(abs(unlist(history[history_count_fields]) -
            round(unlist(history[history_count_fields]))) > 1e-9, na.rm = TRUE)

normalized_facility_duplicates <- history |>
  mutate(name_key = string_key(facility_name)) |>
  count(name_key) |>
  filter(name_key != "", n > 1)

normalized_district_duplicates <- monitor |>
  mutate(district_key = string_key(district)) |>
  count(division, district_key) |>
  filter(district_key != "", n > 1)

district_division_conflicts <- bind_rows(
  monitor |> transmute(source = "Monitoring", division, district = comparison_district),
  history |> transmute(source = "Historical", division, district = comparison_district)
) |>
  distinct() |>
  group_by(district) |>
  filter(n_distinct(division) > 1) |>
  ungroup()

history_month_denominator_failures <- sum(
  abs(history$implied_months_sessions - history$implied_months_participants) > 0.01,
  na.rm = TRUE
)
history_participant_ratio_failures <- sum(
  abs(history$calculated_participants_session - history$average_participants_session) > 0.01,
  na.rm = TRUE
)
history_non_integer_months <- sum(
  abs(history$implied_months_sessions - round(history$implied_months_sessions)) > 0.01,
  na.rm = TRUE
)

monitor_anomaly_rows <- sum(
  monitor$unusual_target_pct |
    monitor$unusual_qa_pct |
    monitor$unusual_sessions_per_facility |
    monitor$unusual_participants_session,
  na.rm = TRUE
)
history_anomaly_rows <- sum(
  history$unusual_monthly_sessions | history$unusual_participants_session,
  na.rm = TRUE
)

weekly_check_failures <- sum(
  abs(monitor$avg_sessions_week - monitor$calculated_sessions_week) > 0.11,
  na.rm = TRUE
)
participant_check_failures <- sum(
  monitor$avg_participants_session != round(monitor$calculated_participants_session),
  na.rm = TRUE
)

validation_checks <- tibble(
  check = c(
    "Numeric parse failures", "Missing required values", "Exact duplicate rows",
    "Normalized-key duplicates", "Non-finite values", "Negative values",
    "Non-positive counts", "Fractional counts", "Invalid percentages",
    "Monitoring formula mismatches", "Historical formula mismatches",
    "Future activation years", "Division conflicts", "Unmatched geographies"
  ),
  failures = c(
    numeric_type_failures,
    sum(is.na(monitor[required_monitor])) + sum(is.na(history[required_history])),
    sum(duplicated(monitor_raw)) + sum(duplicated(history_raw)),
    nrow(normalized_district_duplicates) + nrow(normalized_facility_duplicates),
    non_finite_numeric,
    sum(unlist(monitor[monitor_numeric]) < 0, na.rm = TRUE) +
      sum(unlist(history[history_numeric]) < 0, na.rm = TRUE),
    non_positive_counts,
    non_integer_counts,
    sum(monitor$sessions_vs_target_pct < 0 | monitor$sessions_vs_target_pct > 100) +
      sum(monitor$ccp_quality_assessment_coverage_pct < 0 |
            monitor$ccp_quality_assessment_coverage_pct > 100),
    weekly_check_failures + participant_check_failures,
    history_month_denominator_failures + history_participant_ratio_failures +
      history_non_integer_months,
    sum(history$active_since > as.integer(format(Sys.Date(), "%Y")), na.rm = TRUE),
    nrow(district_division_conflicts),
    sum(district_comparison_all$join_status != "Matched")
  )
)

issue_log <- tibble(
  issue = c(
    "Misspelled header",
    "Location spelling/punctuation variants", "City-vs-district geography",
    "Facility-name formatting/location variants", "Facility categories vs brief",
    "District facility-count disagreement", "Unspecified averaging window",
    "Unusual monitoring values", "Unusual historical values"
  ),
  Frequency = c(
    sum(monitor_raw_names == "General Surgury Sessions"),
    sum(monitor$division_original != monitor$division |
          monitor$district_original != monitor$district) +
      sum(history$division_original != history$division |
            history$district_original != history$district),
    sum(monitor$district != monitor$comparison_district),
    nrow(facility_name_fixes),
    n_distinct(history$facility_type),
    sum(district_comparison_all$facility_count_gap != 0, na.rm = TRUE),
    nrow(history),
    monitor_anomaly_rows,
    history_anomaly_rows
  ),
  Improvements = c(
    "Fixed the spelling in the column name; values stayed the same",
    "Made spelling, spacing, and punctuation consistent",
    "Kept the city rows; used parent districts only when comparing files",
    "Fixed spacing, commas, and clear location spelling errors",
    "Kept all 9 types and noted that the brief says 8",
    "Kept both counts and marked the districts that need checking",
    "Used the given monthly averages and clearly stated the limitation",
    "Kept the values and marked them for checking",
    "Kept the values and marked them for checking"
  )
)

kable(issue_log, align = c("l", "r", "l"))
```

The `r nrow(validation_checks)` basic format and formula checks found no failures. The monitoring file has `r nrow(monitor)` reporting areas. For the cross-file comparison only, I grouped `r sum(monitor$district != monitor$comparison_district)` city rows with their parent districts, leaving `r nrow(district_comparison_all)` comparable districts. Both files cover `r sum(monitor$facilities)` facilities in total, but `r sum(district_comparison_all$facility_count_gap != 0, na.rm = TRUE)` of those districts have different facility counts. I kept both counts because the files do not show which one is current.

**Table: Location names that were fixed**

```{r location-fixes-table}
location_fixes <- bind_rows(
  monitor |>
    filter(district_original != district | division_original != division) |>
    transmute(source = "Monitoring", field = "Location", original = paste(division_original, district_original, sep = " / "), corrected = paste(division, district, sep = " / ") ),
  history |>
    filter(district_original != district | division_original != division) |>
    distinct(division_original, district_original, division, district) |>
    transmute(source = "Historical", field = "Location", original = paste(division_original, district_original, sep = " / "), corrected = paste(division, district, sep = " / ") )
) |>
  distinct()

kable(location_fixes)
```

**Table: Examples of facility names that were fixed**

```{r facility-name-fixes}
kable(head(facility_name_fixes, 10))
```

There are `r nrow(facility_name_fixes)` facility-name fixes in total. They cover line breaks, apostrophes, commas, spacing, and clear district spelling errors. I did not change other words because there is no official facility list in the files.

**Table: Values that should be checked**

```{r anomaly-review}
anomaly_review <- bind_rows(
  monitor |>
    filter(unusual_target_pct) |>
    transmute(source = "Monitoring", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "Sessions vs target %", value = sessions_vs_target_pct),
  monitor |>
    filter(unusual_qa_pct) |>
    transmute(source = "Monitoring", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "QA coverage %", value = ccp_quality_assessment_coverage_pct),
  monitor |>
    filter(unusual_sessions_per_facility) |>
    transmute(source = "Monitoring", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "Sessions per facility", value = total_sessions_quarter / facilities),
  monitor |>
    filter(unusual_participants_session) |>
    transmute(source = "Monitoring", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "Participants per session", value = calculated_participants_session),
  history |>
    filter(unusual_monthly_sessions) |>
    transmute(source = "Historical", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "Average monthly sessions", value = average_monthly_sessions),
  history |>
    filter(unusual_participants_session) |>
    transmute(source = "Historical", row = source_row,
              location = paste(division, district, sep = " / "),
              indicator = "Participants per session", value = calculated_participants_session)
) |>
  arrange(source, indicator, desc(value))

anomaly_display <- bind_rows(
  anomaly_review |> filter(source == "Monitoring"),
  anomaly_review |> filter(source == "Historical") |> slice_head(n = 8)
)
kable(anomaly_display |> mutate(value = round(value, 1)))
```

The full review table contains `r nrow(anomaly_review)` flags. These values look unusual compared with similar rows, but unusual does not always mean wrong. I kept every value and marked it for checking.

**Table: Estimated number of months behind the historical averages**

```{r averaging-window-review}
averaging_window_review <- history |>
  group_by(active_since) |>
  summarise(
    facilities = n(),
    minimum_implied_months = min(implied_months_sessions),
    median_implied_months = median(implied_months_sessions),
    maximum_implied_months = max(implied_months_sessions),
    .groups = "drop"
  ) |>
  mutate(across(ends_with("months"), ~ round(.x, 1)))

kable(averaging_window_review)
```

The session and participant averages use the same number of months for every facility, which is a good sign. However, facilities with the same start year often have very different month counts. Since the file gives only a year, I used the monthly averages as provided and did not guess the missing dates.

# A. What the numbers show

::: {.mainquestion latex=true}
**Main question:** What do the numbers tell us about performance, where do the two files disagree, and where should managers focus?
:::

## Division performance

::: {.subquestion latex=true}
**Question 1:** How do the eight divisions rank when districts with more sessions carry more weight?
:::

**Table: Division ranking from strongest to weakest**

```{r division-results}
division_results <- monitor |>
  group_by(division) |>
  summarise(
    weighted_target_pct = weighted.mean(sessions_vs_target_pct, total_sessions_quarter),
    weighted_qa_pct = weighted.mean(ccp_quality_assessment_coverage_pct, facilities),
    districts = n(),
    facilities = sum(facilities),
    sessions = sum(total_sessions_quarter),
    participants = sum(total_participants),
    trainers = sum(trainers),
    sessions_per_facility = sessions / facilities,
    facilities_per_trainer = facilities / trainers,
    .groups = "drop"
  ) |>
  arrange(desc(weighted_target_pct)) |>
  mutate(performance_rank = row_number())

division_rank_table <- division_results |>
  transmute(
    Rank = performance_rank,
    Division = division,
    Facilities = facilities,
    Sessions = sessions,
    Participants = participants,
    `Weighted target %` = round(weighted_target_pct, 1),
    `Weighted QA %` = round(weighted_qa_pct, 1)
  )

kable(division_rank_table)
```

The target ranking uses session volume as the weight, exactly as requested. This means a district running many sessions has more influence than one running only a few. Mymensingh and Barishal both round to 90.9%; Mymensingh is only slightly higher, so that difference is not meaningful on its own. I could not calculate `total sessions / total target` because district target counts were not provided. For the separate QA measure, I used facility count as the weight because QA coverage is tied more closely to facility coverage; the file does not provide the original QA numerator.

```{r division-chart, echo=FALSE}
ggplot(division_results, aes(x = reorder(division, weighted_target_pct), y = weighted_target_pct)) +
  geom_col(aes(fill = weighted_qa_pct), width = 0.7) +
  coord_flip() +
  geom_text(aes(label = sprintf("%.1f%%", weighted_target_pct)), hjust = -0.1, size = 3) +
  scale_fill_gradient(low = "#d95f02", high = "#1b9e77", name = "QA coverage %") +
  scale_y_continuous(labels = label_percent(scale = 1), expand = expansion(mult = c(0, 0.12))) +
  labs(x = NULL, y = "Sessions compared with target", title = "Division performance and quality checks") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")
```

## Five-district source comparison

::: {.subquestion latex=true}
**Question 2:** For five districts, how does this quarter compare with a three-month estimate based on past monthly averages?
:::

**Table: Reported sessions compared with the three-month estimate**

```{r district-comparison}
comparison_pool <- district_comparison_all |>
  filter(join_status == "Matched", expected_quarter_sessions > 0) |>
  arrange(reported_quarter_sessions) |>
  mutate(volume_quintile = ntile(reported_quarter_sessions, 5))

# Select one middle district from each fifth of the session-volume range.
five_districts <- comparison_pool |>
  group_by(volume_quintile) |>
  arrange(reported_quarter_sessions, .by_group = TRUE) |>
  slice(ceiling(n() / 2)) |>
  ungroup() |>
  mutate(
    absolute_gap_pct = abs(session_gap_pct),
    assessment = case_when(
      absolute_gap_pct > 0.20 ~ "Does not align",
      facility_count_gap != 0 ~ "Totals close; lists differ",
      TRUE ~ "Similar"
    ),
    hypothesis = case_when(
      absolute_gap_pct <= 0.20 & facility_count_gap != 0 ~
        paste0(
          "The totals are close, but monitoring lists ", facilities_monitoring,
          " facilities and the older file lists ", facilities_history,
          ". The close result may be accidental, so check both facility lists and reporting dates."
        ),
      absolute_gap_pct > 0.20 & facility_count_gap > 0 ~
        paste0(
          "Monitoring includes ", facility_count_gap,
          " more facilities, which probably explains part of the higher total. Check the active-facility list and start dates first."
        ),
      absolute_gap_pct > 0.20 & facility_count_gap < 0 ~
        paste0(
          "Monitoring includes ", abs(facility_count_gap),
          " fewer facilities, which probably explains part of the lower total. Check inactive facilities and missing reports first."
        ),
      absolute_gap_pct > 0.20 & facility_count_gap == 0 & session_gap_pct > 0 ~
        "Facility counts match, so the number of facilities does not explain the higher total. Check whether activity grew, the periods differ, or the older average is no longer current.",
      absolute_gap_pct > 0.20 & facility_count_gap == 0 & session_gap_pct < 0 ~
        "Facility counts match, so the number of facilities does not explain the lower total. Check for missed reports, lower activity, and different reporting dates.",
      TRUE ~ "The totals and facility counts are similar; confirm the reporting dates before treating them as a true match."
    )
  )

five_district_table <- five_districts |>
  transmute(
    District = comparison_district,
    `Monitoring sites` = facilities_monitoring,
    `Older-file sites` = facilities_history,
    `This quarter` = reported_quarter_sessions,
    `3-month estimate` = round(expected_quarter_sessions, 1),
    Difference = percent(session_gap_pct, accuracy = 0.1),
    Assessment = assessment
  )

kable(five_district_table)
```

I chose one district from each fifth of the session-volume range, so the sample is not made up only of large or small districts. The estimate is the sum of each facility's average monthly sessions multiplied by three. I used a 20% difference as a review flag, not as a pass mark, because the files do not provide normal month-to-month variation. On that basis, `r sum(five_districts$assessment == "Does not align")` of the five districts do not align. The remaining district has a similar session total but a different facility count, so all five need at least a quick check of reporting dates and active-facility lists before anyone treats the gaps as performance changes.

```{r district-hypotheses, echo=FALSE, results='asis'}
cat("\n")
for (i in seq_len(nrow(five_districts))) {
  row <- five_districts[i, ]
  cat("- **", row$comparison_district, " (", row$assessment, "):** ",
      row$hypothesis, "\n\n", sep = "")
}
```

## Where managers should focus

::: {.subquestion latex=true}
**Question 3:** Which three divisions need the most attention, and what should managers see on a dashboard?
:::

**Table: Three divisions that need the most attention**

```{r priority-analysis}
division_priority <- division_results |>
  mutate(
    target_risk = rescale(100 - weighted_target_pct),
    qa_risk = rescale(100 - weighted_qa_pct),
    productivity_risk = rescale(max(sessions_per_facility) - sessions_per_facility),
    workload_risk = rescale(facilities_per_trainer),
    scale_risk = rescale(facilities),
    priority_score = 100 * (
      0.30 * target_risk + 0.30 * qa_risk +
        0.15 * productivity_risk + 0.15 * workload_risk +
        0.10 * scale_risk
    )
  ) |>
  arrange(desc(priority_score))

priority_three <- head(division_priority, 3)

priority_table <- priority_three |>
  transmute(
    Division = division,
    `Target %` = round(weighted_target_pct, 1),
    `QA %` = round(weighted_qa_pct, 1),
    `Sessions/facility` = round(sessions_per_facility, 1),
    `Facilities/trainer` = round(facilities_per_trainer, 1),
    `Attention score` = round(priority_score, 1)
  )

kable(priority_table)
```

I used five practical signals: missed targets (30%), low QA coverage (30%), fewer sessions per facility (15%), more facilities per trainer (15%), and the number of facilities affected (10%). Program size has only a small weight, so it adds the likely reach of a problem without hiding poor performance in a smaller division. The weights are a management choice, not an official standard, and sessions per facility may also reflect the mix of facility types. The score is therefore a shortlist for discussion, not a final judgment.

**What I would do in the next 30 days**

```{r priority-actions, results='asis'}
priority_actions <- priority_three |>
  mutate(
    reason = case_when(
      division == "Dhaka" ~ sprintf(
        "Largest program (%d facilities), lowest target result (%.1f%%), and %.1f sessions per facility.",
        facilities, weighted_target_pct, sessions_per_facility
      ),
      division == "Chattogram" ~ sprintf(
        "Large program (%d facilities), %.1f%% target result, and low QA coverage (%.1f%%).",
        facilities, weighted_target_pct, weighted_qa_pct
      ),
      division == "Barishal" ~ sprintf(
        "Lowest QA coverage (%.1f%%) and highest trainer load (%.1f facilities per trainer), despite strong session delivery.",
        weighted_qa_pct, facilities_per_trainer
      ),
      TRUE ~ "More than one performance or support measure needs attention."
    ),
    first_action = case_when(
      division == "Dhaka" ~
        "Start with the lowest-target districts, check session totals against registers, and agree a six-week recovery plan with weekly follow-up.",
      division == "Chattogram" ~
        "Start QA follow-up in districts below 60% coverage and pair every check with coaching, an owner, and a due date.",
      division == "Barishal" ~
        "Protect the strong session volume while increasing QA checks and rebalancing or clustering trainer support.",
      TRUE ~ "Review the weakest districts, agree an owner, and track a short action plan."
    )
  ) |>
  transmute(Division = division, `Why now` = reason, `First practical action` = first_action)

cat("\n")
for (i in seq_len(nrow(priority_actions))) {
  row <- priority_actions[i, ]
  cat(
    "- **", row$Division, " — why now:** ", row[["Why now"]],
    " **First action:** ", row[["First practical action"]], "\n\n",
    sep = ""
  )
}
```

## Dashboard for the Implementation Team

| Visual | One decision it should support |
|---|---|
| Division scorecard showing target, QA, sessions per facility, trainer load, and facility count | Which divisions need management attention first, and what is driving the concern? |
| District chart with target result on one axis, QA coverage on the other, and session volume shown by dot size | Is a district's main problem delivery, quality checks, or both? |
| Stacked bars showing session topics by division | Which division needs help restoring a missing or unusually small session topic? |
| Data-check table showing missing reports, unusual values, and differences between the two files | Which figures must be checked before a manager acts on them? |
| Monthly facility activity grid, once monthly facility data are stored | Which facilities have stopped, slowed down, or failed to report for two months? |

The first four views can be built from the current files. The monthly facility grid cannot; it needs facility-level reports to be saved by month from now on. Filters should show only fields that exist for the selected view. Clicking a district should open its supporting figures, the action owner, due date, and latest update.

## Executive summary to the Country Director

::: {.subquestion latex=true}
**Question 4:** What should the Country Director know before the leadership meeting?
:::

```{r executive-summary, echo=FALSE, results='asis'}
strongest <- division_results$division[1]
weakest <- division_results$division[nrow(division_results)]
priority_names <- paste(priority_three$division, collapse = ", ")
misaligned_names <- paste(
  five_districts$comparison_district[five_districts$assessment == "Does not align"],
  collapse = ", "
)

executive_summary <- paste0(
  "Dear Country Director, this quarter's data shows meaningful differences across the eight divisions. ",
  strongest, " ranks first and ", weakest, " ranks last when district target results are weighted by session volume. ",
  "Mymensingh and Barishal are almost tied at 90.9%, so small rank differences should not drive decisions. ",
  "I recommend first attention to ", priority_names, ". Dhaka combines the lowest target result with the largest number of facilities and low sessions per facility. Chattogram has many facilities, below-target delivery, and low QA coverage. Barishal is delivering sessions strongly, but it has the lowest QA coverage and the most facilities per trainer. ",
  "The monitoring file contains 68 reporting areas. After grouping four city areas with their parent districts only for comparison, there are 64 comparable districts. Both files total 658 facilities, yet 59 districts have different facility counts. The brief also says eight facility types, while the older file contains nine. ",
  "In the five-district check, ", misaligned_names, " differ from the three-month estimate by more than 20%. Rangamati's totals are close, but its facility counts still differ. These findings may reflect different facility lists or reporting periods, so they need checking before being treated as performance changes. ",
  "Over the next 30 days, I would confirm the active-facility lists and reporting dates, start a six-week recovery review in Dhaka, close the largest QA gaps in Chattogram and Barishal, and use a monthly dashboard with named owners and due dates."
)

executive_word_count <- str_count(str_squish(executive_summary), "\\S+")
stopifnot(executive_word_count >= 200, executive_word_count <= 250)
cat(executive_summary)
```

*Word count: `r executive_word_count`.*

# B. Monitoring the 15 remote facilities

::: {.mainquestion latex=true}
**Main question:** How should Noora Health monitor 15 hard-to-reach facilities with poor internet, long travel times, and frequent nurse transfers?
:::

## Monitoring approach

::: {.subquestion latex=true}
**Question 1:** What should be tracked, how should data be collected, how often should staff visit, and what are the main risks?
:::

I would keep the routine list to five measures so it is useful without creating too much work for nurses.

| Indicator | Why it matters |
|---|---|
| Sessions delivered against the monthly plan | Shows whether CCP activity is happening at the expected pace. |
| Patients and caregivers reached, reported separately | Shows reach and whether one participant group is being missed. |
| Number of trained nurses still working, including a named backup | Gives an early warning when a transfer could stop delivery. |
| Short session-quality score from a five-item observation checklist | Checks whether sessions are useful, not only frequent. |
| Complete report received by the agreed date | Separates a service problem from a reporting or connectivity problem. |

The facility nurse should record activity in a numbered paper register as the source record, then enter one facility-month total into Noora Health's approved offline form. The form should save on the device, check required fields and unusual values, use a facility-and-month ID to prevent repeat submissions, and upload when a signal is available. [SurveyCTO's documentation](https://docs.surveycto.com/01-getting-started/01-overview/06.collecting-data.html) confirms that its mobile app can collect offline and send data later. A receipt should confirm a successful upload; staff should not resend a form unless no receipt appears. Names or identifiable register photos should not be sent through WhatsApp.

The monitoring officer should make an **initial visit within four to six weeks of launch**, then visit each facility **once per quarter**. An officer covering five facilities would therefore have five planned visits each quarter, usually one or two a month, which leaves time for remote checks and weather delays. Visits can be grouped only when routes and safety allow. During the monthly call, the facility contact should read the aggregate totals from the paper register so the officer can compare them with the submitted form; each visit should include a small source-record check and a live session observation. This is sufficient for routine oversight because direct checks happen every quarter while the monthly calls and warning rules identify problems earlier. One missed deadline should trigger a call; two missed reports, an unexplained zero, a major change, loss of trained staff, or a poor quality score should trigger an earlier visit. This monthly review of completeness, timeliness, unusual changes, and source records follows the practical checks in the [WHO routine health facility data toolkit](https://www.who.int/data/data-collection-tools/health-service-data/toolkit-for-routine-health-information-system-data).

| Expansion-specific risk | Mitigation |
|---|---|
| Trained nurses transfer after 6–9 months | Train a main and backup nurse at launch, require a short handover, and give replacements a remote orientation within two weeks. |
| Poor internet or an app failure causes missing or repeated records | Keep the paper register as the source, use an offline form with a facility-month ID, confirm upload, and compare totals every month. |
| Weather and long travel delay support or create safety risks | Set quarterly routes with backup dates, use a local contact and check-in plan, avoid unsafe solo travel, and use warning signs to decide when an extra visit is worth the journey. |

## Support for the monitoring officer

::: {.subquestion latex=true}
**Question 2:** How would I support one officer covering five facilities without asking them to travel too often?
:::

I would agree a three-month route plan with the officer, provide the travel budget in advance, and protect one day each week for remote review and follow-up. Each facility would name a main and backup contact, so the officer can hold one short monthly call and solve routine problems without travelling. A one-page warning list would show missed deadlines, zero or sharply changing sessions, staff transfers, and open actions; one warning prompts a call, while repeated or serious warnings prompt a visit. For safety, the officer would not make risky river or overnight journeys alone, and every trip would have a weather check, local contact, and check-in/check-out plan. I would meet the officer every two weeks, review a small sample of register totals remotely, arrange cover during leave, and join an occasional visit to coach rather than police.

# C. Fixing the draft monitoring plan

::: {.mainquestion latex=true}
**Main question:** What is weak in the draft plan, what could go wrong, and how should the plan be rewritten?
:::

## Specific weaknesses and risks

::: {.subquestion latex=true}
**Question 1:** What are the main weaknesses, and what risk does each one create?
:::

| Weakness | Concrete risk |
|---|---|
| Only sessions and participants are tracked | High volume can hide poor session quality, missing topics, late reports, or the loss of trained staff. |
| Free-text WhatsApp messages have no fixed definitions, record ID, or receipt | Messages can be missed or repeated, and staff may count the same activity in different ways. |
| Program Managers receive the figures, but the plan does not name a separate person to check the data | Follow-up and data checking can become mixed together, and corrections may not be recorded consistently. |
| The live master spreadsheet is shared with the research team | Extra access, competing edits, and different definitions can create privacy, version, and consent problems. |
| Review happens only each quarter, and only zero sessions are flagged | Missing reports, low quality, sudden changes, and gradual decline may remain open for months. |

## Rewritten plan (4 sentences)

::: {.subquestion latex=true}
**Question 2:** What should the improved plan say?
:::

Each facility will record sessions in a standard register and submit a short offline monthly form by the fifth working day, using fixed definitions and a facility-month ID. The form will check required fields, impossible values, sudden changes, and repeat submissions; the Monitoring Team will compare flagged figures with the source register and record every correction. Program Managers will own follow-up actions, while the Monitoring Team will own data checks and a monthly review of missing reports, zero sessions, low quality, and sharp changes. The research team will receive only a dated, approved copy containing the fields needed for the study, with agreed access and consent rules—not the live program file.

# D. Lessons from the field notes

::: {.mainquestion latex=true}
**Main question:** What common lessons appear in the eight field notes, and what should the program change?
:::

## Themes

::: {.subquestion latex=true}
**Question 1:** How can the eight notes be grouped, and does each theme look isolated or wider than one site?
:::

| Theme | What this tells us | How I would handle it |
|---|---|---|
| Materials do not always fit | Standard content is not being checked often enough against the ward setting and information families already receive. | **Treat as a wider design risk:** two different comments point to poor fit, so check a small sample of facilities before revising the materials. |
| Timing and access need flexibility | One timetable and one audience rule cannot fit daily ward routines, Ramadan, and different family needs. | **Treat as a wider program issue:** three notes point to the same design problem; test better time slots and, where privacy allows, a suitable option for interested family members. |
| Staff changes break continuity | Training depends too heavily on individual nurses instead of a backup and handover system. | **Treat as a wider program issue:** high turnover is already expected and the field note shows its effect, so every facility needs a trained backup and quick replacement support. |
| Problems are reported but not closed | Staff do not have one reliable way to save a problem, see who owns it, and learn what happened next. | **Treat as a wider program issue:** an app failure and the same connectivity problem raised three times without action show both a tool problem and a follow-up problem. |

## Program Learning Note

::: {.subquestion latex=true}
**Question 2:** What short learning note should be shared, and what one process change would help capture feedback better?
:::

```{r learning-note, echo=FALSE, results='asis'}
learning_note <- paste(
  "Recent field feedback shows that several small design and support gaps are reducing session quality.",
  "Materials do not always match the ward setting, and some content repeats information families have already seen.",
  "Attendance changes with ward routines and Ramadan, while current participation rules may leave out interested family members.",
  "Nurse transfers also leave new staff without enough support.",
  "At the same time, staff are losing data in the app and do not know whether problems raised on calls are being acted on.",
  "I recommend one shared offline feedback and action log linked to the monthly monitoring form.",
  "Each issue should record the site, theme, urgency, owner, due date, and closure note.",
  "The Monitoring and Implementation teams should review the log together each month and send the final action back to the person who raised it."
)

learning_word_count <- str_count(str_squish(learning_note), "\\S+")
stopifnot(learning_word_count >= 100, learning_word_count <= 150)
cat(learning_note)
```

*Word count: `r learning_word_count`.*

# Exported data workbook

The cleaned source data, comparison tables, results, and data checks are saved in `Updated_Data.xlsx`.

```{r export-workbook}
# Each named data frame becomes one worksheet in the Excel file.
export_sheets <- list(
  Clean_Monitoring = monitor,
  Clean_Facilities = history,
  Division_Results = division_results,
  District_Comparison = district_comparison_all,
  Five_Districts = five_districts,
  Priority_Divisions = division_priority,
  Priority_Actions = priority_actions,
  Issue_Log = issue_log,
  Validation_Checks = validation_checks,
  Location_Fixes = location_fixes,
  Facility_Name_Fixes = facility_name_fixes,
  Anomaly_Review = anomaly_review,
  Averaging_Windows = averaging_window_review
)

write_xlsx(export_sheets, "Updated_Data.xlsx")
```

# What the data cannot tell us

- The files do not say which quarter was reported or the exact period behind the older averages. The quarter estimates are only a rough comparison, not a prediction.
- Most district facility counts differ between the two files. I could not settle these differences because the files do not share a fixed facility ID or dated facility list.
- City rows were grouped with their parent districts only when comparing the two files. The original monitoring locations were kept.
- I kept the rounded averages from the source files. Session and participant totals were added again from the detailed columns so the maths could be checked.
- Every cleaning step and calculation is in this R Markdown file. No cleaned value was changed by hand outside the code.

# References used for the practical recommendations

- World Health Organization. [Toolkit for Routine Health Information Systems data](https://www.who.int/data/data-collection-tools/health-service-data/toolkit-for-routine-health-information-system-data).
- SurveyCTO. [Collecting data online and offline](https://docs.surveycto.com/01-getting-started/01-overview/06.collecting-data.html).
